---
title: "PETRUS_VAL: Inter-observer Validation of Ultrasound Reading Quality"
author: "Clinical Study Report"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
    theme: united
    highlight: tango
    code_folding: hide
  pdf_document:
    toc: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 8,
  fig.height = 6,
  dpi = 300
)

# Load required packages
library(tidyverse)
library(irr)
library(psych)
library(knitr)
library(kableExtra)

# Source custom functions
source("../R/functions.R")
```

# Executive Summary

This report presents the results of the inter-observer validation study for ultrasound reading quality. The study evaluates the agreement between multiple observers in assessing ultrasound images.

**Key Findings:**

- [To be completed after analysis]

# Introduction

## Study Objectives

The primary objectives of this study are:

1. To assess inter-observer agreement in ultrasound reading quality
2. To identify variables with high and low agreement
3. To quantify reliability using appropriate statistical measures
4. To provide recommendations for standardization and training

## Methods

### Statistical Analyses

The following statistical methods were employed:

- **Percentage Agreement**: Simple agreement percentage between observers
- **Cohen's Kappa** (κ): For categorical variables (unweighted)
- **Weighted Kappa**: For ordinal variables (quadratic weights)
- **Intraclass Correlation Coefficient (ICC)**: For continuous measurements (ICC(2,1) two-way random effects, single rater)
- **Bland-Altman Analysis**: For assessing agreement and bias in continuous measurements

### Interpretation Guidelines

**Cohen's Kappa / Weighted Kappa:**

- < 0.00: Poor (less than chance agreement)
- 0.00 - 0.20: Slight
- 0.21 - 0.40: Fair
- 0.41 - 0.60: Moderate
- 0.61 - 0.80: Substantial
- 0.81 - 1.00: Almost Perfect

**Intraclass Correlation Coefficient (ICC):**

- < 0.50: Poor
- 0.50 - 0.75: Moderate
- 0.75 - 0.90: Good
- > 0.90: Excellent

# Data Overview

```{r load-data, eval=FALSE}
# Load preprocessed data
# Uncomment and modify path as needed
# paired_data <- readRDS("../data/processed/paired_data.rds")
# analysis_results <- readRDS("../data/processed/analysis_results.rds")
```

## Sample Characteristics

```{r sample-characteristics, eval=FALSE}
# Display sample size and characteristics
# cat("Total number of paired observations:", nrow(paired_data), "\n")
```

## Data Quality

```{r data-quality, eval=FALSE}
# Report on missing data and completeness
```

# Results

## Overall Agreement Statistics

```{r summary-table, eval=FALSE}
# Load and display summary table
# summary_results <- read_csv("../data/processed/summary_results.csv")
# 
# kable(summary_results, 
#       caption = "Summary of Inter-observer Agreement Statistics",
#       digits = 3) %>%
#   kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
#                 full_width = FALSE)
```

## Categorical Variables

### Variable 1: [Variable Name]

```{r categorical-var1, eval=FALSE}
# Display results for categorical variable
# result <- analysis_results[["Variable Name"]]
# 
# cat("Cohen's Kappa:", round(result$kappa$value, 3), "\n")
# cat("Interpretation:", interpret_kappa(result$kappa$value), "\n")
# 
# # Display confusion matrix
# kable(result$confusion_matrix,
#       caption = "Confusion Matrix: Observer 1 vs Observer 2") %>%
#   kable_styling(bootstrap_options = c("striped", "hover"))
```

## Continuous Variables

### Variable 2: [Variable Name]

```{r continuous-var1, eval=FALSE}
# Display results for continuous variable
# result <- analysis_results[["Variable Name"]]
# 
# cat("ICC(2,1):", round(result$icc$results$ICC[2], 3), "\n")
# cat("Interpretation:", interpret_icc(result$icc$results$ICC[2]), "\n")
# cat("Pearson Correlation:", round(result$correlation$estimate, 3), "\n")
```

#### Bland-Altman Plot

```{r bland-altman-var1, eval=FALSE, fig.cap="Bland-Altman plot showing agreement between observers"}
# Display Bland-Altman plot
# print(result$bland_altman_plot)
```

#### Scatter Plot

```{r scatter-var1, eval=FALSE, fig.cap="Scatter plot comparing observer measurements"}
# Display scatter plot
# print(result$scatter_plot)
```

# Discussion

## Strengths and Limitations

**Strengths:**

- [To be completed]

**Limitations:**

- [To be completed]

## Clinical Implications

[To be completed based on results]

## Recommendations

Based on the inter-observer agreement analysis:

1. [Recommendation 1]
2. [Recommendation 2]
3. [Recommendation 3]

# Conclusions

[To be completed based on results]

# References

1. Cohen J. A coefficient of agreement for nominal scales. Educ Psychol Meas. 1960;20(1):37-46.
2. Landis JR, Koch GG. The measurement of observer agreement for categorical data. Biometrics. 1977;33(1):159-174.
3. Shrout PE, Fleiss JL. Intraclass correlations: uses in assessing rater reliability. Psychol Bull. 1979;86(2):420-428.
4. Bland JM, Altman DG. Statistical methods for assessing agreement between two methods of clinical measurement. Lancet. 1986;1(8476):307-310.
5. Koo TK, Li MY. A Guideline of Selecting and Reporting Intraclass Correlation Coefficients for Reliability Research. J Chiropr Med. 2016;15(2):155-163.

# Appendix

## Session Information

```{r session-info}
sessionInfo()
```

## Data Processing Details

```{r processing-details, eval=FALSE}
# Document any data processing steps
```
